---
title: "football_data"
author: "Frank Agyemang"
date: "2022-11-27"
output: html_document
---

Loading libraries
```{r}
library(datasets)
library(tidyr)
library(ggplot2)
library(skimr)
library(dplyr)
library(plyr)
library(linearModel)
library(GGally)
library(corrplot)
library(ggpubr)
library(broom)
#library(tidyverse)
library(cowplot)
library(car)

```

#Reading in the dataset
```{r}
football_data <- read.csv("players_20.csv")
skim(football_data)
```

#The skim function in R gives a summary of the data. The difference between the skim function and summary stems from the fact that the skim
#function provides more details about the data than the summary function.
#Looking at the output from the skim function tells me whether I may need to do some data clearing with some of the variables of interest.
# Note that it groups the variables by data type and that is helpful with the type of analysis we make.
```{r}
summary(football_data)
```

```{r}
football_data_sub <- football_data %>%
  select(short_name,age,height_cm, weight_kg,overall, wage_eur,player_positions,attacking_crossing,attacking_heading_accuracy,attacking_short_passing,attacking_volleys,attacking_finishing)

```

# Analyzing the distribution of the variables selected
```{r}
ggplot(football_data_sub, aes(age)) + 
  geom_bar(fill = "yellow" , color = "black")+
  scale_x_binned()
```


# From the plot it can be observed that majority of the players are around the age 24years.
# We could use this information to determine the probability of a player been called to the world given his age.

```{r}
age_frequency <- count(football_data_sub, "age") %>%
  mutate(probability = age_frequency$freq/length(football_data_sub$age))
age_frequency

par(mar = c(1, 1, 1,1))
plot(age_frequency$age,age_frequency$probability, col = "black", xlab = "age",ylab = "prob")

```

# #Analyzing the distribution of the height variable
# ggplot(football_data_sub, aes(height_cm,fill=height_cm)) +
#   geom_bar()+
#   scale_fill_gradient2(low = "red",high = "yellow",mid = "green")


# Checking whether there are relationship between pairs of variables
#height and weight

```{r}
par(mar = c(1, 1, 1,1))
norm_height <- log(football_data_sub$height_cm)
norm_weight <- log(football_data_sub$weight_kg)
plot(norm_height,norm_weight, col="blue")
abline(lm(norm_weight~norm_height))
correlation <- cor(norm_height,norm_weight)
```

#exploratory analysis of wage_eur

```{r}
min(football_data_sub$wage_eur)
max(football_data_sub$wage_eur)
wage <- football_data_sub$wage_eur
wage <- wage[wage > 0]
wage
min(wage)
max(wage)
var(wage)
sd(wage)
mean(wage)

football_data_sub1 <- subset(football_data_sub, select = -c(short_name,player_positions))
pairs(football_data_sub1, pch = 18, col = "black")
hist(wage,
     col="grey",
     border="black",
     prob = TRUE)

lines(density(wage,adjust = 2),
      lwd = 1,
      col = "black")

```

#The density plot shows that the wage of players may be skewed.

```{r, warning=FALSE}
football_data_sub %>% 
  select(age:wage_eur,attacking_crossing:attacking_finishing) %>% 
  ggpairs()
```


#Does a player's Age and Weight affect a player's attacking finishing

```{r}
attacker_finish <- football_data_sub%>%
  select(age,weight_kg,attacking_finishing)

```


#corelation plots

```{r}
plot(attacker_finish$age, attacker_finish$attacking_finishing, pch = 19, col = "blue")

abline(lm(attacker_finish$attacking_finishing ~ attacker_finish$age), col = "black", lwd = 3)
cor(attacker_finish$age, attacker_finish$attacking_finishing)
```

# The plot and correlation coefficient shows that age has a positive correlation with attackers 
#finishing
# This is plausible because older players in soccer tend to have more experience 
# and thus have better finishing that young footballers.

```{r}
plot(attacker_finish$weight_kg, attacker_finish$attacking_finishing, pch = 19, col = "blue")
abline(lm(attacker_finish$attacking_finishing ~ attacker_finish$weight_kg), col = "black", lwd = 3)
cor(attacker_finish$weight_kg, attacker_finish$attacking_finishing)
```

#The second correlation plots also shows that weight has a negative correlation with an 
# attackers finishing .
#This is also plausible because players who weigh more are not able to run faster and
# hence are not better attackers. The heavier a player becomes to more negative it affects his finishing.


```{r}
attacking_finish_model_1<-lm(attacker_finish$attacking_finishing ~ attacker_finish$age + attacker_finish$weight_kg, data = attacker_finish)
```

Checking for interaction using *

```{r}
attacking_finish_model_2<-lm(attacker_finish$attacking_finishing ~ attacker_finish$age * attacker_finish$weight_kg, data = attacker_finish)

summary(attacking_finish_model_2)
```

Now lets's check for collinearity using vif()

The values in our result is greater than 5. Therefore there is a problem. We have to center and redo the model.

```{r}
vif(attacking_finish_model_2)
```

Redoing the model
Standardize the data

```{r}
stand <- attacker_finish %>% 
  mutate(stAge=age - mean(age)) %>% 
  mutate(stWeight_kg=weight_kg-mean(weight_kg))
```
Making a new model

```{r}
attacking_finish_model_3  <- lm(stand$attacking_finishing ~ stand$stAge * stand$stWeight_kg, data = stand)

summary(attacking_finish_model_3)
```


