---
title: "ML_FIFA WORLD CUP"
Author: "John Mensah"
format: html
editor: visual
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
library(randomForest)
library(rsample) 
library(spatialRF)
library(randomForestExplainer)
#library(Rcpp)
library(recipes)
library(tidymodels)
fifa_data <- read.csv("players_20.csv")
player_rate <- fifa_data %>% select( ,11,32:37)
player_rate <- na.omit(player_rate)
```

### 1. Which of the six compositions contribute largely to the OVR of the players?

```{r}
### this random forest analysis takes not less than 30min to run
### I have to do this because the rshiny app is not working for a reason I do not know. 
set.seed(1234)
ctrl <- trainControl(method = "cv",  number = 10)
co_split <- initial_split(player_rate, prop = .7)
co_train <- training(co_split)
nrow(co_train)
co_test <- testing(co_split)
nrow(co_test)



data_recipe<-recipe(overall   ~ ., 
                data = co_train) %>%
  step_impute_knn(all_predictors()) %>%
step_center(all_numeric_predictors()) %>%
  step_scale(all_numeric_predictors())
# process the traing set/ prepare recipe(non-cv)
data_prep <-data_recipe %>% prep(training = co_train, retain = TRUE)


rf_model<-rand_forest(trees = 100, mtry = 6, mode = "regression") %>% set_engine("randomForest",importance=T, localImp = T, ) %>% fit(overall   ~ ., data = juice(data_prep))

impt_frame<-measure_importance(rf_model$fit)
#impt_frame[order(-impt_frame$node_purity_increase), , drop = FALSE]

md_frame <- min_depth_distribution(rf_model$fit)

# IncNodePurity

#imp.rf <- measure_importance(rf_model$fit) %>%
imp.rf <- impt_frame %>% 
  as.data.frame() %>%
 rownames_to_column("Variable") %>%
  arrange(node_purity_increase)
#imp.rf


plot_min_depth_distribution(md_frame, mean_sample = "top_trees")

imp.rf %>%
  mutate(Variable = factor(variable, levels = reorder(variable, node_purity_increase))) %>%
  ggplot(aes(x = node_purity_increase, y = reorder(variable, +node_purity_increase), fill = variable)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(aspect.ratio = 1,
        legend.position = "none",
        axis.text.y = element_text(size = 16,
                                   face = "bold")
  ) +
  scale_y_discrete("") +
  scale_x_continuous("Inc Node Purity") + labs(caption = 
"Fig.2. Random forest analysis of the six main statistics that 
composite the overall rating of the FIFA World Cup players. 
physic = physicality of the players") +
  theme(plot.caption = element_text(hjust=0))

#Identifying potential interactions
vars<- important_variables(measure_importance(rf_model$fit), k = 6)



forest <- randomForest::randomForest(overall   ~ ., data = bake(data_prep, new_data = co_train), localImp = TRUE, importance=T, ntree = 100, mtry = 6, type= "regression")

```

```{r}
players_rate <- fifa_data %>% select( ,32:37)
library(ggplot2)
#install.packages("ggcorrplot")
library(ggcorrplot)

  
# Computing correlation matrix
player_rate <- na.omit(player_rate)
correlation_matrix <- round(cor(player_rate),3)

  
# Computing correlation matrix with p-values
corrp.mat <- cor_pmat(player_rate)
  
# Adding correlation significance level
ggcorrplot(correlation_matrix, hc.order =TRUE, type ="lower", 
           p.mat = corrp.mat, lab = T) + theme_classic() + ylab("") + xlab("")
```

### Interpretation

Rating of football players is one of the most anticipated exercise in the game of association football. The players, fans, investors, and any other person involved in the game look forward to the overall rating of the players. This is very important because it tells the overall performance of the player for the season under review, and also influences the price, contract and market demand of a player. Six main factors composite the overall rating of the football players, but whether these factors differ in their overall contribution and how much they differ is not known. Also, how much variation in the overall rating is explained by these factors is quite not clear even though it has been noted that the greatest proportion of the overall rating is dependent on six main factors: speed, shooting, passing, dribbling, defending and physicality of the players.

In order to evaluate the total variation and individual contribution of these six main factors to the overall rating of the FIFA World Cup players, we fitted a Random Forest model to the data. The random forest model revealed that, the six main factors considered in the overall rating of the players explains a total variation of 96.4% in the players' overall rating. This is much expected because the six factors considered are the main factors found on the FIFA Ultimate Team (FUT) card (Fig. 1).The FUT card contains information (including the player's name, photo, nation's flag, name, and club's badge) and statistics (the players rating on pace, shooting, dribbling, defense, passing, and physicality) of a real life footballer and is categorised based on the player's type and quality. These is statistics is commonly known to many as the only determinant of the players' OVR, but it is not always the case. In some cases, it has been noted that the players international reputation can influence the OVR by adding onto it.

Figure 2 measure the contribution of each of the six statistics on the FUT card on the players' OVR. We determined this based on the nodes of purity of a fitted regression trees. The concept of the node of purity in the regression tree is such that, the fitted model is a one big bushy forest (here a forest of 500 trees) and the various explanatory variables are factors that sustain the forest. Now, the forest is trimmed down by removing these explanatory variables that sustain the forest, and it done in a such a way that if one variable is removed how much of the forest would be reduced. The variable that will lead to the highest effect on the forest is classified as the one with the most increased node of purity (in other words, the most important variable). In fig 2, the dribbling skills of the players holds the highest node of purity, followed by defending skills and the passing skills, respectively. While the players pace was the least. The result in fig 2 reveal that the six statistics on the FUT card contribute differently in terms of their proportions to the OVR of the players and thus, the ability of a player to dribble or defend increases the chance of having a higher OVR. In other words, players with high dribbling or defending skills are more likely to have higher OVR compared to players with high pace or physicality.

*Blas M. Benito (2021). spatialRF: Easy Spatial Regression with Random Forest. R package version 1.1.0. doi: 10.5281/zenodo.4745208. url:<https://blasbenito.github.io/spatialRF/>*

```{r}
spatialRF::plot_training_df(
  data = player_rate,
  dependent.variable.name = "overall",
  predictor.variable.names = colnames(player_rate)[2:7],
  ncol = 3,
  point.color = viridis::viridis(100, option = "F"),
  line.color = "gray30",
  method = "gam"
  )


```

### Interpretation

Rating of football players is one of the most anticipated exercise in the game of association football. The players, fans, investors, and any other person involved in the game look forward to the overall rating of the players. This is very important because it tells the overall performance of the player for the season under review, and also influences the price, contract and market demand of a player. Six main factors composite the overall rating of the football players, but whether these factors differ in their overall contribution and how much they differ is not known. Also, how much variation in the overall rating is explained by these factors is quite not clear even though it has been noted that the greatest proportion of the overall rating is dependent on six main factors: speed, shooting, passing, dribbling, defending and physicality of the players.

In order to evaluate the total variation and individual contribution of these six main factors to the overall rating of the FIFA World Cup players, we fitted a Random Forest model to the data. The random forest model revealed that, the six main factors considered in the overall rating of the players explains a total variation of 96.4% in the players' overall rating. This is much expected because the six factors considered are the main factors found on the FIFA Ultimate Team (FUT) card (Fig. 1).The FUT card contains information (including the player's name, photo, nation's flag, name, and club's badge) and statistics (the players rating on pace, shooting, dribbling, defense, passing, and physicality) of a real life footballer and is categorised based on the player's type and quality. These is statistics is commonly known to many as the only determinant of the players' OVR, but it is not always the case. In some cases, it has been noted that the players international reputation can influence the OVR by adding onto it^3.

Figure 2 measure the contribution of each of the six statistics on the FUT card on the players' OVR. We determined this based on the nodes of purity of a fitted regression trees. The concept of the node of purity in the regression tree is such that, the fitted model is a one big bushy forest (here a forest of 500 trees) and the various explanatory variables are factors that sustain the forest. Now, the forest is trimmed down by removing these explanatory variables that sustain the forest, and it done in a such a way that if one variable is removed how much of the forest would be reduced. The variable that will lead to the highest effect on the forest is classified as the one with the most increased node of purity (in other words, the most important variable). In fig 2, the dribbling skills of the players holds the highest node of purity, followed by defending skills and the passing skills, respectively. While the players pace was the least. The result in fig 2 reveal that the six statistics on the FUT card contribute differently in terms of their proportions to the OVR of the players and thus, the ability of a player to dribble or defend increases the chance of having a higher OVR. In other words, players with high dribbling or defending skills are more likely to have higher OVR compared to players with high pace or physicality.

## 2. Does the preferred foot of the players matter when it comes to players performance?

```{r}
library(ggstatsplot)
library(tidyverse)
fifa_data <- read.csv("players_20.csv")

PF <- ggbetweenstats(
  data = fifa_data,
  x = preferred_foot,
  y = overall,
  pairwise.comparisons = F,
  bf.message = TRUE
) + labs(
    x = "Preferred foot",
    y = "Overall rating"
  ) +
  theme(
    axis.ticks = element_blank(),
    axis.line = element_line(colour = "grey50"),
    panel.grid = element_line(color = "#b4aea9"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(linetype = "dashed"),
    panel.background = element_rect(fill = "#fbf9f4", color = "#fbf9f4"),
    plot.background = element_rect(fill = "#fbf9f4", color = "#fbf9f4")
  ) + labs(caption = 
"Fig.3. Preferred foot and the overall rating of the FIFA World Cup players") +
  theme(plot.caption = element_text(hjust=0))

PF

```

There has been a generational debate about left and right footed footballers, which of them are most talented or play very well. Generally, we do not observe as many left footed footballers as right footed footballers but the few left footed players are usually known to be very good. In fig 3, we compared the overall rating of the left and right preferred foot of the players to see which of them has the highest OVR and if the difference is significant. In total, there were 4318 left footed players and 13960 right footed players. We found that, the left footed payers have significantly higher OVR (mean = 66.88) than the right footed players (mean = 66.11) (Fig 3).


### 3. Does the international reputation of the players corresponds to the OVR of the players?

```{r}
library(PMCMRplus)
ggbetweenstats(
  data = fifa_data,
  x = international_reputation,
  y = overall,
  pairwise.comparisons = F,
  bf.message = TRUE
) + labs(
    x = "International reputation (Star)",
    y = "Overall rating"
  ) +
  theme(
    axis.ticks = element_blank(),
    axis.line = element_line(colour = "grey50"),
    panel.grid = element_line(color = "#b4aea9"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(linetype = "dashed"),
    panel.background = element_rect(fill = "#fbf9f4", color = "#fbf9f4"),
    plot.background = element_rect(fill = "#fbf9f4", color = "#fbf9f4")
  ) + labs(caption = 
"Fig.4 International reputation rating and the overall rating of the FIFA World Cup players") +
  theme(plot.caption = element_text(hjust=0))
```

The international reputation of the players is rated on the basis of how known/long they have played for their country. The is very important rating as it help boost the OVR of the players. In Fig 4, we found that higher international reputation significantly increases the players OVR. Playing at the national teams comes with a lot of sacrifices, and many players goes to the extent to switch nationality in order to earn a call up to play at the international football competitions. Our result explains why many players are willing to make so many sacrifices with hard work in order to play for their countries. International reputation help improves not only the players OVR but also the player market demand and wages.

### 4. How much effect do the WFS rating have on the players OVR?

```{r}
library(rstantools)
ggbetweenstats(
  data = fifa_data,
  x = weak_foot,
  y = overall,
  pairwise.comparisons = F,
  bf.message = F
) + labs(
    x = "Weak foot Skill (star)",
    y = "Overall rating"
  ) +
  theme(
    axis.ticks = element_blank(),
    axis.line = element_line(colour = "grey50"),
    panel.grid = element_line(color = "#b4aea9"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(linetype = "dashed"),
    panel.background = element_rect(fill = "#fbf9f4", color = "#fbf9f4"), plot.background = element_rect(fill = "#fbf9f4", color = "#fbf9f4"),
    panel.background = element_rect(fill = "#fbf9f4", color = "#fbf9f4"),
    plot.background = element_rect(fill = "#fbf9f4", color = "#fbf9f4")
  ) + labs(caption = 
"Fig.5. Weak foot skill rating and the overall rating of the FIFA World Cup players") +
  theme(plot.caption = element_text(hjust=0))

```

The ability of a football player to equally use the weak as the preferred foot is also very important determinant of players rating. In Fig 5, we found that players that are able to use both their preferred and weak foot equally have higher rating compared to those with less ability to use both foot equally. Being able to use both foot is an important skill to possess as a football player because it makes one unpredictable by the opponent and stand a very high chance of winning possession of the ball. Such players are also very versatile in the field and are able to play in positions that might require only left or right footed players.

### 5. Does the players SM matter in the OVR of the players?

```{r}
ggbetweenstats(
  data = fifa_data,
  x = skill_moves,
  y = overall,
  pairwise.comparisons = F,
  bf.message = TRUE
) + labs(
    x = "Skill moves (star)",
    y = "Overall rating"
  ) +
  theme(
    axis.ticks = element_blank(),
    axis.line = element_line(colour = "grey50"),
    panel.grid = element_line(color = "#b4aea9"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(linetype = "dashed"),
    panel.background = element_rect(fill = "#fbf9f4", color = "#fbf9f4"),
    plot.background = element_rect(fill = "#fbf9f4", color = "#fbf9f4")
  ) + labs(caption = 
  )  + labs(caption = 
"Fig.6. Skill move rating and the overall rating of the FIFA World Cup players") +
  theme(plot.caption = element_text(hjust=0))
```

Skill moves are also very important in the game pf association football. Players with low skill moves are most likely to loose possession of the ball or have less dribbling efficiency. Fig 6 also showed that higher skill moves significantly increases the players OVR.

### 6. How does the combination of the players working rating (i.e., attacking working rate/defensive working rate) contribute to the OVR?

```{r}
replace <- function(x){
  y <- x
  for(i in 1:length(x)){
    if(x[i] == "High/High"){
      y[i] <- "H/H"
    } else if(x[i] == "High/Medium"){
      y[i] <- "H/M"
    }else if(x[i] == "High/Low"){
      y[i] <- "H/L"
    } else if(x[i] == "Medium/Medium"){
      y[i] <- "M/M"
    }else if(x[i] == "Medium/Low"){
      y[i] <- "M/L"
    }else if(x[i] == "Medium/High"){
      y[i] <- "M/H"
    }else if(x[i] == "Low/Low"){
      y[i] <- "L/L"
    }else if(x[i] == "Low/Medium"){
      y[i] <- "L/M"
    }else if(x[i] == "Low/High"){
      y[i] <- "L/H"
    }else {
      y[i] <- x[i]
    }
      
  } 
  return(y)
}

fifa_data$work_rate <- replace(fifa_data$work_rate)

ggbetweenstats(
  data = fifa_data,
  x = work_rate,
  y = overall,
  pairwise.comparisons = F,
  palette = "Spectral",
  bf.message = F
) + labs(
    x = "Working rate",
    y = "Overall rating"
  ) +
  theme(
    axis.ticks = element_blank(),
    axis.line = element_line(colour = "grey50"),
    panel.grid = element_line(color = "#b4aea9"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(linetype = "dashed"),
    panel.background = element_rect(fill = "#fbf9f4", color = "#fbf9f4"),
    plot.background = element_rect(fill = "#fbf9f4", color = "#fbf9f4")
  ) + theme(
  plot.caption = element_text(hjust = 0)
  ) + labs(caption = "Fig. 7. Differences in overall rating among the nine different categories of working rate 
of the FIFA World Cup players. H = High, M = Medium, L = Low. The rating
was done by comparing the offensive working rate to the defensive working rate of the players") +
  theme(plot.caption = element_text(hjust=0))
  + theme(
  plot.caption = element_text(hjust = 0)
  ) + labs(caption = "Fig. 7. Differences in overall rating among the nine different categories of working rate 
of the FIFA World Cup players. H = High, M = Medium, L = Low. The rating
was done by comparing the offensive working rate to the defensive working rate of the players") + theme(plot.caption = element_text(hjust=0))

#View(paletteer::palettes_d_names)
```

Work rate of the players is very important in football game. It help increases the chance of winning ball possession and decreasing the possibility of conceding goals. The work rate of the players on a scale of high, medium, and low assessed based on the attacking (offensive) vs defensive working rate. This help measure if the player has the same offensive working rate as the defensive working rate and vice versa. Knowledge of players working rate is very key to deciding the formation or strategy of the team for a particular game. In fig 7, we compared the OVR of the FIFA players with respect to their working rate. The purpose is to understand the disparity in offensive vs defensive working rate in the OVR of the players. We found that, the OVR differs significantly among the nine different categories of working rate. On the average, players with high offensive and defensive working rate has the highest OVR (mean = 70.41), followed by players with medium offensive and high defensive working rate (mean = 68.37), high offensive and low defensive working rate (mean = 68.10), and high offensive and medium defensive working rate (mean = 68.06), respectively. While players with low offensive and defensive working rate have the least OVR (mean = 61.86).


### Just some test analysis- not necessary

```{r}
install.packages("ggfortify")
library(plotly)
library(ggfortify)
pc_data <- fifa_data %>% select(,16,17,25,11,32:37)
pc_data <- na.omit(pc_data)
pc_data<-  pc_data %>% filter(!team_position == "")
#pc_data %>% remove_rownames %>% column_to_rownames(var="team_position")

rate.pca <- prcomp(pc_data[,c(5:10)], center = T,scale. = T)

components <- data.frame(rate.pca[["x"]])

components <- cbind(components, pc_data$team_position)

components$PC2 <- -components$PC2
explained_variance <- summary(rate.pca)[["sdev"]]
explained_variance <- explained_variance[1:2]
comp <- rate.pca[["rotation"]]
comp[,'PC2'] <- - comp[,'PC2']
loadings <- comp
for (i in seq(explained_variance)){
  loadings[,i] <- comp[,i] * explained_variance[i]
}

features = c(unique(pc_data$team_position))

fig <- plot_ly(components, x = ~PC1, y = ~PC2, color = ~pc_data$team_position, type = 'scatter', mode = 'markers') %>%
  layout(
    legend=list(title=list(text='color')),
    plot_bgcolor = "#e5ecf6",
    xaxis = list(
      title = "0"),
    yaxis = list(
      title = "1"))
for (i in seq(4)){
  fig <- fig %>%
    add_segments(x = 0, xend = loadings[i, 1], y = 0, yend = loadings[i, 2], line = list(color = 'black'),inherit = FALSE, showlegend = FALSE) %>%
    add_annotations(x=loadings[i, 1], y=loadings[i, 2], ax = 0, ay = 0,text = features[i], xanchor = 'center', yanchor= 'bottom')
}

autoplot(rate.pca, data = pc_data, colour = "team_position", loadings = T, loadings.colour = 'blue', loadings.label = TRUE, loadings.label.size = 4, loadings.label.colour='black') + theme_linedraw()


autoplot(rate.pca, data = pc_data, colour = "international_reputation", loadings = T, loadings.colour = 'blue', loadings.label = TRUE, loadings.label.size = 4, loadings.label.colour='red') + theme_linedraw()

autoplot(rate.pca, data = pc_data, colour = "preferred_foot", loadings = T, loadings.colour = 'blue', loadings.label = TRUE, loadings.label.size = 4, loadings.label.colour='black') + theme_linedraw()
```

## Analysis by Frank Agyemang

## What factors contribute to the salary of a player?
Sports is one of the attractive field in our world today because of the numerous benefits such as the recognition and the money. Like many other known sports, football is no exception. It is known that sportsmen are among the highly paid professionals in our world today. Nevertheless, an interesting debate that pops up most of the time is on what should be the right salary for a footballer. For some football lovers certain players deserve the amount they take whilst others do not. In this work, we try to determine some factors that could influence the wage of a player(s) who participate in the FIFA world cup. The variables we consider will be Pace, Shooting, Passing, Dribbling, Defending , Physic,International reputation, overall, wage_eur.

We could also use the information to determine the approximate budget for the world cup.

The salary of a player has been a debate among football lovers for a long time.
```{r}
football_data <- read.csv("players_20.csv")
summary(football_data)
```

selecting the variables of interest
```{r}
library(dplyr)

football_data_sort <- football_data %>%
  select(pace, shooting, passing, dribbling, defending, physic,international_reputation, overall, wage_eur)
# football_data_sort
```

Now, I see that there are some missing data in some of the columns and therefore I have to remove them
```{r}
library(tidyverse)
football_data_drop <- na.omit(football_data_sort)
football_data_drop
# length(football_data_drop$pace)
```


Plots of the variables
```{r}
library(ggplot2)
# plot(football_data_drop, pch = 9)

## Correlation plots
library(corrplot)
corrplot(corr = cor(football_data_drop),
         addCoef.col = "black",
         number.cex = 0.5,
         number.digits =1 ,
         diag = F,
         bg = "magenta" ,
         outline = "black",
         addgrid.col = "white",
         mar = c(1,1,1,1))
```
We make correlation plots to determine the level of association various pairs of explanatory variables and response variables. In the plot shown, variables that are negatively correlated have a color of brown and variables that are positively correlated have a color of light blue. It can be seen that International reputation and the overall rating of a player in the FIFA world has the strongest positive correlation with wage followed by the passing and dribbling skills of a player. 
Looking at the correlation plots there are also some level of association between some pairs of explanatory variables. For example, the overall rating of a player is strong and positively correlated with passing, dribbling and shooting. 


## We now perform simple linear regression using variables that had a correlation coefficient greater than or equal to 5. That is, Overall Reputation and International reputation.
```{r}
library(dplyr)
# football_data_drop
library(linearModel)
int_rep <- lm(football_data_drop$wage_eur ~ football_data_drop$international_reputation,data = football_data_drop)
anova(int_rep)
# int_rep

plot(int_rep,pch=7)
# hist(football_data_drop$international_reputation)
library(moments)
skewness(football_data_drop$international_reputation, na.rm = TRUE) 
```
We can see that the residual plots show that there is unequal spread of data points and in the Normal Q-Q plots, the data points gradually move far away from the fitted line. Hence, the data does not follow a normal distribution  and this can also be confirmed by the histogram plot which is also skewed. The value for skewness is far greater than 1 and that is worrying. Therefore we need to transform the data.


Residual Plots 
```{r}
library(tidyverse)
library(broom)
theme_set(theme_classic())
int_rep.diag.metrics <- augment(int_rep)
head(int_rep.diag.metrics)
```

```{r}
library(ggplot2)
ggplot(int_rep.diag.metrics, aes(football_data_drop$international_reputation,football_data_drop$wage_eur )) +
  geom_point() +
  stat_smooth(method = lm, se = FALSE) +
  geom_segment(aes(xend = football_data_drop$international_reputation, yend = .fitted), color = "red", size = 0.3)
```

```{r}
library(dplyr)
# football_data_drop
library(linearModel)
overall_model <- lm(football_data_drop$wage_eur ~ football_data_drop$overall,data = football_data_drop)
anova(int_rep)
overall_model

plot(overall_model,pch=7)
#hist(football_data_drop$international_reputation)
library(moments)
skewness(football_data_drop$overall, na.rm = TRUE)
```


From the above plots, I want to transform all the variables so that they follow a normal distribution.
```{r}
speed_transformed <- qplot(x = log10(pace + 1),data = football_data_drop)
int_rep_transformed <- qplot(x = sqrt(international_reputation),data = football_data_drop)
overall_transformed <- qplot(x = log10(overall + 1),data = football_data_drop)
passing_transformed <- qplot(x = log10(passing + 1),data = football_data_drop)
dribbling_transformed <- qplot(x = log10(dribbling + 1),data = football_data_drop)
wage_transformed <- qplot(x=log10(wage_eur + 1),data = football_data_drop)

# speed_transformed
# int_rep_transformed
# overall_transformed
# passing_transformed
# dribbling_transformed
# wage_transformed
```


```{r}
library(dplyr)
pace_transformed <- log10(football_data_drop$pace + 1)
int_rep_transformed <-  sqrt(football_data_drop$international_reputation)
overall_transformed <- log10(football_data_drop$overall + 1)
passing_transformed <- log10(football_data_drop$passing + 1)
dribbling_transformed <- log10(football_data_drop$dribbling + 1)
wage_transformed <- log10(football_data_drop$wage_eur + 1)

football_data_drop <- football_data_drop%>%
   mutate(pace_transformed = pace_transformed, int_rep_transformed = int_rep_transformed,
         overall_transformed = overall_transformed, passing_transformed = passing_transformed,
         dribbling_transformed = dribbling_transformed, wage_transformed = wage_transformed)
 
```

We now re-run the models 
```{r}
library(linearModel)
overall_model <- lm(football_data_drop$wage_transformed ~ football_data_drop$overall_transformed,data = football_data_drop)
anova(int_rep)
overall_model

pace_model <- lm(football_data_drop$wage_transformed ~ football_data_drop$pace_transformed,data = football_data_drop)
anova(pace_model)
pace_model

int_rep_model <- lm(football_data_drop$wage_transformed ~ football_data_drop$int_rep_transformed,data = football_data_drop)
anova(int_rep_model)
int_rep_model

# pace_model <- lm(football_data_drop$wage_transformed ~ football_data_drop$speed_transformed,data = football_data_drop)
# anova(pace_model)
# pace_model

dribbling_model <- lm(football_data_drop$wage_transformed ~ football_data_drop$dribbling_transformed,data = football_data_drop)
anova(dribbling_model)
dribbling_model
```

We look at the residual plots again
```{r}
library(tidyverse)
library(broom)
theme_set(theme_classic())
int_rep_model.diag.metrics <- augment(int_rep_model)
overall_model.diag.metrics <- augment(overall_model)
pace_model.diag.metrics <- augment(pace_model)
dribbling_model.diag.metrics <- augment(dribbling_model)

head(int_rep_model.diag.metrics)
head(overall_model.diag.metrics)
head(pace_model.diag.metrics)
head(dribbling_model.diag.metrics)
```

```{r}
library(ggplot2)
ggplot(int_rep_model.diag.metrics, aes(football_data_drop$int_rep_transformed,football_data_drop$wage_transformed )) +
  geom_point() +
  stat_smooth(method = lm, se = FALSE) +
  geom_segment(aes(xend = football_data_drop$int_rep_transformed, yend = .fitted), color = "red", size = 0.3)

ggplot(overall_model.diag.metrics, aes(football_data_drop$overall_transformed,football_data_drop$wage_transformed )) +
  geom_point() +
  stat_smooth(method = lm, se = FALSE) +
  geom_segment(aes(xend = football_data_drop$overall_transformed, yend = .fitted), color = "red", size = 0.3)
```


```{r}
ggplot(pace_model.diag.metrics, aes(football_data_drop$pace_transformed,football_data_drop$wage_transformed )) +
  geom_point() +
  stat_smooth(method = lm, se = FALSE) +
  geom_segment(aes(xend = football_data_drop$pace_transformed, yend = .fitted), color = "red", size = 0.3)
```



Now i make a regression model taking into consideration all the selected factors at the same time.
```{r}
library(linearModel)
multiple_reg <- lm(football_data_drop$wage_transformed ~ football_data_drop$pace_transformed +football_data_drop$int_rep_transformed + football_data_drop$overall_transformed + football_data_drop$pace_transformed, data = football_data_drop)
anova(multiple_reg)
multiple_reg
```

Residual plots
```{r}
res <- resid(multiple_reg)
plot(fitted(multiple_reg), res,pch=1)
abline(0,0)
```

```{r}
qqnorm(res)

#add a straight diagonal line to the plot
qqline(res)
```


## Analysis by Christian Uwineza

# Players rating against players value

```{r}
# players with a rating greater than 90
rating_90<- football_data %>% filter(overall>=90)
# platers with a rating between 85 and 90
rating_85_90<- football_data %>% filter(overall>=85 & overall<90)

# value vs rating

ggplot(data = football_data)+
  geom_line(
    aes(x= value_eur, y= overall)
  )+
xlab("Players' value (€)")

``` 

```{r}

# bar plot comparing the ovall rating of players positions
ggplot(data = football_data)+
  geom_bar(
    aes(x= team_position, y= overall, fill= team_position),
    stat = "identity",
    width= 0.5
  )+
  theme_bw()+
  ggtitle("Overal rating per players position")+
  xlab("Team position")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
   scale_y_continuous(
    name = "Rating (%)",
    limits = c(0,100),
    breaks = seq(0,100, by= 10))
ggsave(path = "/Users/cb/Desktop/project-cfj_r-is-cool", file = "rating_position.jpeg", width = 12, height =8, dpi=700)


```
The bar graph above shows different positions that exist in football and how each position is generally rated. 

### Teams paying their players more than others. 
We took a closer look at the players salary and grouped all players from the same team together then added all their salaries together to come up with a sum of the totals teams are paying their players. We used the geom_col inside the gglot2 to make a nice graph with the top teams paying a lot to their players. 


```{r}
football_data %>% 
  group_by(club) %>% 
  filter(value_eur>=3.15e+07) %>% 
  summarise(total_value= sum(as.integer(value_eur), na.rm=TRUE)) %>% 
  ggplot(aes(reorder(club, total_value), total_value, fill= total_value))+
  geom_col(show.legend = TRUE)+
  theme_bw()+
  coord_flip()+
  scale_y_continuous(
    name = "Football clubs values (€)",
    limits = c(0,700000000),
    breaks = seq(0,700000000, by= 200000000))+
  theme(axis.title.y = element_blank())+
  ggtitle("Most expensive football clubs")+
      theme(axis.text=element_text(size=12),
        axis.title=element_text(size=15,face="bold"))+
  theme(legend.text = element_text(size = 12)) +
  theme(plot.title = element_text(size = 20, face = "bold")) +
  theme(text = element_text(size = 17))+
labs(caption = 
"Fig.8. Top football clubs paying more money to their players") +
  theme(plot.caption = element_text(hjust=0))+
    theme(
    axis.line = element_line(colour = "grey50"),
    panel.grid = element_line(color = "#b4aea9"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_line(linetype = "dashed"),
    panel.grid.major.y = element_blank(),
    panel.background = element_rect(fill = "#fbf9f4", color = "#fbf9f4"),
    plot.background = element_rect(fill = "#fbf9f4", color = "#fbf9f4")
  )

ggsave(path = "/Users/cb/Desktop/project-cfj_r-is-cool", file = "Expensive_teams.jpeg", width = 12, height =8, dpi=700)

  
 #labels = c("0 €", "1 Billion €","2 Billion €","3 Billion €","4 Billion €","5 Billion €","6 Billion €","7 Billion €","8 Billion €"
```
From the graph above, FC Barcelona, Real Madrid, Manchester City, FC Bayern Munchen and Liverpool are the top five teams that are paying a lot of money their players. The top two teams are from Spain. We also have two teams from England and one team from Germany. 


# Comparison between players within same team (Overall and poetential rating)

```{r}
football_data %>% 
  filter(club=="FC Barcelona") %>% 
  select(short_name, overall, potential) %>% 
  arrange(-overall) %>% 
  gather(Variable, Exp, -short_name) %>% 
  ggplot(aes(short_name, Exp, fill=Variable))+
  geom_col(position = "dodge")+
  geom_text(aes(label=Exp), position = position_dodge(width = 0.7), vjust=-0.5, size = 2.5) + xlab("Name of players") + ylab("Rating of players (%)") +
  scale_fill_manual(values = c("#065535","#0000b3"))+
  theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ggtitle("FC Balcelona's players rating")+
    theme(axis.text=element_text(size=12),
        axis.title=element_text(size=15,face="bold"))+
  theme(legend.text = element_text(size = 15)) +
  theme(plot.title = element_text(size = 20, face = "bold")) +
  theme(text = element_text(size = 17))+
labs(caption = 
"Fig.9. FC Barcelona's players rating") +
  theme(plot.caption = element_text(hjust=0))+
    theme(
    axis.line = element_line(colour = "grey50"),
    panel.grid = element_line(color = "#b4aea9"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.background = element_rect(fill = "#fbf9f4", color = "#fbf9f4"),
    plot.background = element_rect(fill = "#fbf9f4", color = "#fbf9f4")
  )

ggsave(path = "/Users/cb/Desktop/project-cfj_r-is-cool", file = "Barcelona's_rating.jpeg", width = 16, height =8, dpi=700)


```
Since FC Barcelona is the top paying football club, we take a look deep into their players rating in the graph above. This team has at least 11 players with an overall rating greater than or equal to 85.
