---
title: "Stat 850 Project Description"
author: "Christian Uwineza"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(digits = 3)
```


## Instructions

Each member of your team should modify this document in some way and push their modifications to the repository in a separate commit. This will ensure that you have set your repository up in a way that ensures all group members are working with the same repository. 

Note that you must compile your readme (this document) for it to properly display as part of your github repository. 

Once you have received feedback on your project proposal (via Canvas) you may alter this README so that it describes your final project instead of the project proposal. 

## Data Set
Provide a link to your dataset here, and a description of the variables in the data set.

https://www.kaggle.com/datasets/thedevastator/fifa-world-cup-anomaly-detection-in-player-ratin?select=players_20.csv


## Potential Topics to Explore Using the Data Set
- Players rating versus players value
- Top football clubs player's ratings (overall and potential rating)
- comparison between players 
- Teams with high economic value

## Group Members
List all of the project contributors here.
John Mensah
Frank Agyemang
Christian Uwineza


## Dr. Vanderplas' Exploration of the Data

Ideally, this would have been something you provided as part of your data explanations :).

```{r, message = F, warning = F}
library(tidyverse)
library(skimr)
library(dplyr)
football_data <- read.csv("players_20.csv")

```


Variables: 

- player_url
- short_name
- long_name
- age
- Date of Birth
- height_cm
- weight_kg
- nationality
- club
- overall
- potential
- value_eur
- wage_eur
- player_positions
- preferred_foot
- weak_foot
- skill_moves
- work_rate
- body_type
- player_traits
- attacking_crossing
- attacking_finishing
- attacking_volleys
- skill_dribbling
- skill_curve
- movement_balance
- movement_reactions
- power_jumping
- power_stamina
- overall


# Players rating against platers value

```{r}
# players with a rating greater than 90
rating_90<- football_data %>% filter(overall>=90)
# platers with a rating between 85 and 90
rating_85_90<- football_data %>% filter(overall>=85 & overall<90)

# value vs rating

ggplot(data = football_data)+
  geom_line(
    aes(x= value_eur, y= overall)
  )+
xlab("Players' value (€)")

``` 

```{r}

# bar plot comparing the ovall rating of players positions
ggplot(data = football_data)+
  geom_bar(
    aes(x= team_position, y= overall, fill= team_position),
    stat = "identity",
    width= 0.5
  )+
  theme_bw()+
  ggtitle("Overal rating per players position")+
  xlab("Team position")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
   scale_y_continuous(
    name = "Rating (%)",
    limits = c(0,100),
    breaks = seq(0,100, by= 10))
ggsave(path = "/Users/cb/Desktop/project-cfj_r-is-cool", file = "rating_position.jpeg", width = 12, height =8, dpi=700)


```
The bar graph above shows different positions that exist in football and how each position is generally rated. 

### Teams paying their players more than others. 
We took a closer look at the players salary and grouped all players from the same team together then added all their salaries together to come up with a sum of the totals teams are paying their players. We used the geom_col inside the gglot2 to make a nice graph with the top teams paying a lot to their players. 


```{r}
football_data %>% 
  group_by(club) %>% 
  filter(value_eur>=3.15e+07) %>% 
  summarise(total_value= sum(as.integer(value_eur), na.rm=TRUE)) %>% 
  ggplot(aes(reorder(club, total_value), total_value, fill= total_value))+
  geom_col(show.legend = TRUE)+
  theme_bw()+
  coord_flip()+
  scale_y_continuous(
    name = "Football clubs values (€)",
    limits = c(0,700000000),
    breaks = seq(0,700000000, by= 200000000))+
  theme(axis.title.y = element_blank())+
  ggtitle("Most expensive football clubs")

ggsave(path = "/Users/cb/Desktop/project-cfj_r-is-cool", file = "Expensive_teams.jpeg", width = 12, height =8, dpi=700)

  
 #labels = c("0 €", "1 Billion €","2 Billion €","3 Billion €","4 Billion €","5 Billion €","6 Billion €","7 Billion €","8 Billion €"
```
From the graph above, FC Barcelona, Real Madrid, Manchester City, FC Bayern Munchen and Liverpool are the top five teams that are paying a lot of money their players. The top two teams are from Spain. We also have two teams from England and one team from Germany. 


# Comparison between players within same team (Overall and poetential rating)

```{r}
football_data %>% 
  filter(club=="FC Barcelona") %>% 
  select(short_name, overall, potential) %>% 
  arrange(-overall) %>% 
  gather(Variable, Exp, -short_name) %>% 
  ggplot(aes(short_name, Exp, fill=Variable))+
  geom_col(position = "dodge")+
  geom_text(aes(label=Exp), position = position_dodge(width = 0.7), vjust=-0.5, size = 2.5) + xlab("Name of players") + ylab("Rating of players (%)") +
  scale_fill_manual(values = c("#065535","#0000b3"))+
  theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ggtitle("FC Balcelona's players rating")+
    theme(axis.text=element_text(size=12),
        axis.title=element_text(size=15,face="bold"))+
  theme(legend.text = element_text(size = 15)) +
  theme(plot.title = element_text(size = 20, face = "bold")) +
  theme(text = element_text(size = 17))  

ggsave(path = "/Users/cb/Desktop/project-cfj_r-is-cool", file = "Barcelona's_rating.jpeg", width = 16, height =8, dpi=700)


```
Since FC Barcelona is the top paying football club, we take a look deep into their players rating in the graph above. This team has at least 11 players with an overall rating greater than or equal to 85. 


