---
title: "Stat 850 Project Description"
author: "Agyemang, Frank, Kusi"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(digits = 3)
```

```{r load-sas-libraries, echo = F, message = F, warning = F}

saspath <- "/usr/local/SASHome/SASFoundation/9.4/bin/sas_en" # Your SAS path goes here
sasopts <- "-ls 75"
knitr::opts_chunk$set(engine.path=list(sas=saspath, saslog=saspath, sashtml=saspath),
                      engine.opts=list(sas=sasopts, saslog=sasopts, sashtml=sasopts), 
                      comment=NA)
library(SASmarkdown)
```

## Instructions

Each member of your team should modify this document in some way and push their modifications to the repository in a separate commit. This will ensure that you have set your repository up in a way that ensures all group members are working with the same repository. 

Note that you must compile your readme (this document) for it to properly display as part of your github repository. 

Once you have received feedback on your project proposal (via Canvas) you may alter this README so that it describes your final project instead of the project proposal. 

## Data Set
Provide a link to your dataset here, and a description of the variables in the data set.

https://prism.oregonstate.edu/explorer/tmp/TsuDeeu02CXPNDjf/PRISM_ppt_tmean_stable_4km_198501_201012_45.0000_-123.0000.csv

> This link doesn't work for me. - Dr. Vanderplas

Drought Data: This data was obtained from https://spei.csic.es. It is a website that grants free access to global drought data. The drought was measured as a standardized precipitation evapotranspiration index (SPEI) - which means it accounts for the combined effect of temperature and rainfall. The dataset span from January 1950 to September 2020. However, we will only filter out data from January 1990 to December 2009, which is the period that covers the monitoring period of our plant population data. The drought data was obtained for the specific study site (Arapaho Prairie Reserve) using the GPS coordinate of the site(Lat: 41.492004    Long: -101.84592). The link to the drought data is here: https://soton.eead.csic.es/spei/task/done/SPEI_41.25_-101.75.csv

> This seems to have date and SPEI data, which is 2 variables. 

> This link also doesn't work for me - Dr. Vanderplas

## Potential Topics to Explore Using the Data Set
Describe avenues you might explore using the data
Rainfall and Temperature: This data was obtained from the Prism Climate Group website (https://prism.oregonstate.edu). It is not in very good shape. But the data, generally, has three variables: date, mean monthly temperature (tmean; degree Celcius), and total monthly rainfall (ppt; mm). The date span from January 1985 to December 2010.

## Group Members
List all of the project contributors here.
John Mensah
Frank Agyemang
Christian Uwineza
