---
title: "Stat 850 Project Description"
author: "Agyemang, Frank, Kusi"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(digits = 3)
```

```{r load-sas-libraries, echo = F, message = F, warning = F}

saspath <- "/usr/local/SASHome/SASFoundation/9.4/bin/sas_en" # Your SAS path goes here
sasopts <- "-ls 75"
knitr::opts_chunk$set(engine.path=list(sas=saspath, saslog=saspath, sashtml=saspath),
                      engine.opts=list(sas=sasopts, saslog=sasopts, sashtml=sasopts), 
                      comment=NA)
library(SASmarkdown)
```

## Instructions

Each member of your team should modify this document in some way and push their modifications to the repository in a separate commit. This will ensure that you have set your repository up in a way that ensures all group members are working with the same repository. 

Note that you must compile your readme (this document) for it to properly display as part of your github repository. 

Once you have received feedback on your project proposal (via Canvas) you may alter this README so that it describes your final project instead of the project proposal. 

## Data Set
Provide a link to your dataset here, and a description of the variables in the data set.

https://www.kaggle.com/datasets/thedevastator/fifa-world-cup-anomaly-detection-in-player-ratin?select=players_20.csv


Football Data: This data was obtained from https://kaggle.com. It is a website that grants free access to data on vast topics.  The data comprises of several variables that can be used for interesting analysis in FIFA World Cup. The Fifa World Cup is the biggest stage in international football, and the players that take part in it are some of the best in the world. But what goes into making a great World Cup player? Is it the wage of a player? Is it their individual ratings for specific skillsets?  Or something unrelated to the variables considered in this dataset?

This dataset contains information on players overall ratings, as well as their individual ratings for specific skillsets. This data can be used to probe into so many aspects related to the World Cup and its players, including anomaly detection for player ratings.


## Potential Topics to Explore Using the Data Set
Describe avenues you might explore using the data



## Group Members
List all of the project contributors here.
John Mensah
Frank Agyemang
Christian Uwineza


## Dr. Vanderplas' Exploration of the Data

Ideally, this would have been something you provided as part of your data explanations :).

```{r, message = F, warning = F}
library(tidyverse)
library(skimr)
transitions <- read_csv("Transitions.csv") %>% select(-1) # remove row names
weather <- read_csv("weather.data.csv") %>% select(-1) # remove row names
fate <- read_csv("fate.trans.csv") %>% select(-1) # remove row names
```

### Weather Data
```{r}
skim(weather)
```

Variables: 

- player_url
- short_name
- long_name
- age
- Date of Birth
- height_cm
- weight_kg
- nationality
- club
- overall
- potential
- value_eur
- wage_eur
- player_positions
- preferred_foot
- weak_foot
- skill_moves
- work_rate
- body_type
- player_traits
- attacking_crossing
- attacking_finishing
- attacking_volleys
- skill_dribbling
- skill_curve
- movement_balance
- movement_reactions
- power_jumping
- power_stamina

### Fate.trans.csv

Where does this data come from? How do you download it (what inputs are required?)
What does each variable mean?

```{r}
skim(fate)
```

Variables:

- Genet
- Ramet
- Year
- Month1
- stage2
- PrevStage
- PrevYear
- NextStage
- NextYear
- firstYear
- firstStage
- lastYear
- lastStage

This data seems to be coded in a way that might be very redundant? Can you explain why it is formatted this way and what you might have to do to make it tidy? What units are the variables in? How does it connect to your other datasets?

### transitions.csv


Where does this data come from? How do you download it (what inputs are required?)
What does each variable mean?  What units are the variables in? How does it connect to your other datasets?

Variables: 

- Year
- Month1
- Dead
- F
- MR
- PD
- S
- SR
- UK

The labels here seem to be similar to the stages in the previous dataset, but what is stored in the columns? Counts? I can't figure this out.

### Overall assessment

At the moment, I don't have any idea about

- the provenance of this data (what do I have to do to get the same files you have?)
- what the variables mean
- how the datasets are connected
- what analysis/visualizations you plan to do with the data

Because of this, I can't figure out whether your data meet the bare minimum requirements for the project. You seem to have enough cases and variables, but without knowing what the variables mean and what the structure is, I can't guarantee that you have information that will give you enough to work with to meet my expectations on this project. 

Please consider making an appointment with me to discuss so that we can work this out - I want to help  and I want you to do a project on data that's interesting to you, but I also have to make sure I'm setting you up for success. 

### What I still need:

- Variable explanations, how the datasets are connected

- What, specifically, are you going to do with this data? Are you going to make a shiny applet to show how drought conditions are related to SPEI? That seems like an obvious thing, but is there something not-so-obvious that I'm missing? Are you going to map this spatially? Create animations to show the extent of drought in an area over time? I really can't figure out what the goal is here.
